Draft Final Report: Measles risk assessment, modelling, and benefit--cost analyses for New Zealand

### 2014-11-07 draftfinalreport.tex
hmmm

The Prioritized Ethnicities for cases are:  European; Maori; Pacific Peoples; Asian; Middle Eastern/Latin American/African (MLA); Other Ethnicity; Residual Categories. For the analyses in this report only the first five are used, as these categories cover the overwhelming number of cases, with only 1.9\% (22/1137) of cases having no Prioritized Ethnicity (see \emph{None}, Table~\ref{table:none}).

@@

The numbers of cases per category and population sizes for the complete data set from 2007-2014 can be seen in Table~\ref{table:percap}. Subsequent regression analyses (not shown) also suggested that the MLA category was over-- or underrepresented in per capita rates given the very small sample sizes for this classification (Figure~\ref{fig:percap}, Table~\ref{table:percap}), leading to very large standard error in regression analyses. However, there are numerous issues with the data for MLA ethnicity category above and beyond the small population sizes (see Table~\ref{table:percap}), thus creating issues for estimating the denominator data for this group (University of Otago, personal communication). Thus, we removed this grouping for our subsequent analyses and are left with Asian, European, Maori and Pacific as Prioritized Ethnicities. This left us with 1102/1115 (99\%) of the measles cases with Prioritized Ethnicity recorded from 2007, and 1102/1137 (97\%) of all measles cases recorded since 2007 (Table~\ref{table:percap}).

\begin{figure}
\begin{center}
\includegraphics{draftfinalreport-010}
\end{center}
\caption{Per capita rates of measles infections for specific age categories (0--2, 3--5, 6--17, 18--24, and 24+ years) and New Zealand deprivation indices (NZDep, 1--5 and 6--10) for 2007--2014, broken down by ethnicity (Table~\ref{table:percap} for details)}
\label{fig:percap}
\end{figure}

@@

The distribution of the meases cases per category used in the regression analyses are in Figure~\ref{fig:histcasecat}.
\begin{figure}
\begin{center}
\includegraphics{draftfinalreport-014}
\end{center}
\caption{Distribution of measles cases per category (Table~\ref{table:percap}) used in the final regression model (Equation~\ref{eq:reg})}
\label{fig:histcasecat}
\end{figure}

The predicted values from the regression model plotted against the reported cases are shown in Figure~\ref{fig:predict}, and the residuals are shown in Figure~\ref{fig:resid}.
\begin{figure}
\begin{center}
\includegraphics{draftfinalreport-015}
\end{center}
\caption{Regression model (Equation~\ref{eq:reg}) predictions plotted against the case data (Table~\ref{table:percap})}
\label{fig:predict}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics{draftfinalreport-017}
\end{center}
\caption{Histogram of residuals from the regression model (Equation~\ref{eq:reg})}
\label{fig:resid}
\end{figure}

The significance of the different predictor variables can be seen in the ANOVA results (Table~\ref{table:anova}).
\par
\vspace{5mm} %5mm vertical space
\begin{table}
% latex table generated in R 3.0.2 by xtable 1.7-3 package
% Sat Nov 08 01:33:38 2014
\begin{tabular}{lrrrrrr}
  \hline
 & Df & Deviance & Resid. Df & Resid. Dev & F & Pr($>$F) \\ 
  \hline
NULL &  &  & 39 & 1453.62 &  &  \\ 
  Age & 4 & 1304.20 & 35 & 149.43 & 183.58 & 0.0000 \\ 
  Ethnicity & 3 & 20.00 & 32 & 129.43 & 3.75 & 0.0285 \\ 
  NZDep & 1 & 10.70 & 31 & 118.74 & 6.02 & 0.0239 \\ 
  Age:Ethnicity & 12 & 81.91 & 19 & 36.83 & 3.84 & 0.0045 \\ 
   \hline
\end{tabular}\caption{Significance of different predictor variables for measles risk factors from 2007-2014, estimated using Equation~\ref{eq:reg} and a quasipoisson error structure}
\label{table:anova}
\end{table}
\vspace{5mm} %5mm vertical space

A summary of the regression model (Equation~\ref{eq:reg}) with the individual effects and the statistical support for the estimated coefficients can be seen in Table~\ref{table:regression}.

\vspace{5mm} %5mm vertical space
\begin{table}
% latex table generated in R 3.0.2 by xtable 1.7-3 package
% Sat Nov 08 01:33:38 2014
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
  \hline
(Intercept) & -6.4674 & 0.1148 & -56.32 & 0.0000 \\ 
  Age3-5 & -0.9129 & 0.2054 & -4.44 & 0.0003 \\ 
  Age6-17 & -0.7618 & 0.1343 & -5.67 & 0.0000 \\ 
  Age18-24 & -1.5050 & 0.1937 & -7.77 & 0.0000 \\ 
  Age25+ & -2.9507 & 0.1608 & -18.35 & 0.0000 \\ 
  EthnicityAsian & 0.0530 & 0.3250 & 0.16 & 0.8721 \\ 
  EthnicityMaori & 0.2124 & 0.1994 & 1.07 & 0.2999 \\ 
  EthnicityPacific & 1.1603 & 0.2043 & 5.68 & 0.0000 \\ 
  NZDep6-10 & -0.2119 & 0.0855 & -2.48 & 0.0227 \\ 
  Age3-5:EthnicityAsian & -2.0315 & 1.3827 & -1.47 & 0.1581 \\ 
  Age6-17:EthnicityAsian & -1.0074 & 0.4717 & -2.14 & 0.0459 \\ 
  Age18-24:EthnicityAsian & -0.8311 & 0.5941 & -1.40 & 0.1779 \\ 
  Age25+:EthnicityAsian & 0.4211 & 0.4433 & 0.95 & 0.3541 \\ 
  Age3-5:EthnicityMaori & -0.3864 & 0.4096 & -0.94 & 0.3573 \\ 
  Age6-17:EthnicityMaori & -0.0855 & 0.2469 & -0.35 & 0.7328 \\ 
  Age18-24:EthnicityMaori & -0.5828 & 0.4484 & -1.30 & 0.2092 \\ 
  Age25+:EthnicityMaori & -1.0482 & 0.5242 & -2.00 & 0.0600 \\ 
  Age3-5:EthnicityPacific & -2.1316 & 0.8139 & -2.62 & 0.0169 \\ 
  Age6-17:EthnicityPacific & -1.4541 & 0.3347 & -4.35 & 0.0003 \\ 
  Age18-24:EthnicityPacific & -0.9827 & 0.5129 & -1.92 & 0.0705 \\ 
  Age25+:EthnicityPacific & -0.6127 & 0.4367 & -1.40 & 0.1767 \\ 
   \hline
\end{tabular}\caption{Summary of the regression model results for measles risk factors for cases from 2007-2014}
\label{table:regression}
\end{table}
\vspace{5mm} %5mm vertical space

\par Apart from over-representation of some MLA categories discussed above and not included here, the results of the regression model suggest that age is a strong predictor of being a measles case. Indeed, all age categories are significantly less likely to be measles cases compared to 0-2 year olds, and the likelihood generally decreases with age (see also Figure~\ref{fig:ageinyears}).

The majority of measles cases are among people of European ancestry (Table~\ref{table:percap} and Figure~\ref{fig:ageinyears}), however, people of Pacific origin are also over-represented as measles cases on a per capita basis ($\beta$ = 1.16, standard error (SE) = 0.2, p--value < 0.0001). Those people in NZDep levels 6-10 are under-represented ($\beta$ = -0.21, SE = 0.085, p--value = 0.02), and there are some other age:ethnicity classes that are significantly less represented in the data compared to Europeans in those ages classes, particularly in the 6--17 age classes. In later measles outbreaks (since 2007) there has been a shift in the distribution of ages infected. The very young (<2 years old) are still most likely to be infected, but school aged children and older teenagers are more likely to be represented then the under tens (Figures~\ref{fig:ageinyears} and ~\ref{fig:ageandvac}, Table~\ref{table:sero}, and section~\ref{sub:popim}). This pattern suggests that improving vaccination coverage in the young is reducing the burden of measles in those age categories (Table~\ref{table:percap}). Interestingly the regression results suggest risk of measles in the 6--17 year age category was greater for Europeans and Maori.

@@

The majority of the na\"{\i}ve among the general New Zealand population are in their first years of life (Figure~\ref{fig:naive}). However, the distribution of na\"{\i}ve at a national level shows that the recent MMR vaccination schemes are reducing the proportions of na\"{\i}ve population in the 3--5 year old age group to greater levels than in older young people  (Figure~\ref{fig:naive}). Plotting the breakdown of these figures by DHB clearly shows that the greatest numbers of na\"{\i}ve people are in DHBs with larger urban areas (Figure~\ref{fig:naive_map}). The distribution of the numbers of na\"{\i}ve and the total na\"{\i}ve populations per DHB, assuming national immunisation and immunity rates are representative, are given in the following figures:
\begin{itemize}[noitemsep,nolistsep]
\item Northland: Figure~\ref{fig:Northland}
\item Waitemata: Figure~\ref{fig:Waitemata}
\item Auckland: Figure~\ref{fig:Auckland}
\item Counties Manukau: Figure~\ref{fig:Counties_Manukau}
\item Waikato: Figure~\ref{fig:Waikato}
\item Lakes: Figure~\ref{fig:Lakes}
\item Bay of Plenty: Figure~\ref{fig:BayofPlenty}
\item Tairawhiti: Figure~\ref{fig:Tairawhiti}
\item Taranaki: Figure~\ref{fig:Taranaki}
\item Hawke's Bay: Figure~\ref{fig:HawkesBay}
\item Whanganui: Figure~\ref{fig:Whanganui}
\item Mid-Central: Figure~\ref{fig:Midcentral}
\item Hutt: Figure~\ref{fig:Hutt}
\item Capital and Coast: Figure~\ref{fig:CapitalandCoast}
\item Wairarapa: Figure~\ref{fig:Wairarapa}
\item Nelson Marlborough: Figure~\ref{fig:NelsonMarlborough}
\item West Coast: Figure~\ref{fig:WestCoast}
\item Canterbury: Figure~\ref{fig:Canterbury}
\item South Canterbury: Figure~\ref{fig:SouthCanterbury}
\item Southern: Figure~\ref{fig:Southern}
\end{itemize}

\begin{figure}
\begin{center}
\includegraphics{draftfinalreport-023}
\end{center}
\caption{New Zealand population by age and estimated numbers of na\"{\i}ve people in each age class using national immunity data (Table~\ref{table:sero})}
\label{fig:naive}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=1.2\textwidth]{naive_map.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:naive_map}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb1.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Northland District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:Northland}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb2.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Waitemata District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:Waitemata}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb3.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Auckland District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:Auckland}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb4.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Counties Manukau District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:Counties_Manukau}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb5.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Waikato District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:Waikato}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb6.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Lakes District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:Lakes}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb7.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Bay of Plenty District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:BayofPlenty}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb8.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Tairawhiti District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:Tairawhiti}
\end{figure}

\begin{figure}[H]
  \begin{center}
    \includegraphics[width=0.6\textwidth]{dhb9.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Taranaki District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:Taranaki}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb10.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Hawke's Bay District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:HawkesBay}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb11.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Whanganui District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:Whanganui}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb12.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Midcentral District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:Midcentral}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb13.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Hutt District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:Hutt}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb14.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Capital and Coast District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:CapitalandCoast}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb15.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Wairarapa District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:Wairarapa}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb16.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Nelson Marlborough District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:NelsonMarlborough}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb17.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, West Coast District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:WestCoast}
\end{figure}

\begin{figure}[H]
   \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb18.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Canterbury District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:Canterbury}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb19.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, South Canterbury District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:SouthCanterbury}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb20.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Southern District Health Board, using national immunity data (Table~\ref{table:sero})}
     \label{fig:Southern}
\end{figure}

@@

Though the global incidence of measles is declining, in recent years that decline has slowed (Figure~\ref{fig:trendincidence}) and immigration rates to New Zealand have risen (Figure~\ref{fig:travel}). This suggests that the risk of measles importation could increase, though further analyses are require to understand the interaction between these variables.  Of note, however, is the clear seasonality in immigration and travel (Figure~\ref{fig:travel}).  This seasonality suggests that there may be period of increased risk of measles importation, though again the interactions with seasonal measles transmission from the nations of origin will be an important factor in determining the risk of measles importation. Of interest is the asynchrony, or lack of phase, between peak non-New Zealander immigration and New Zealanders travelling. The Christmas summer period being an exception to this, when there is substantial New Zealander travel and combined these lead to a very sharp peak in travel rates in December (Figure~\ref{fig:travel}).

>>>>

The Prioritized Ethnicities for cases are:  European; Maori; Pacific Peoples; Asian; Middle Eastern/Latin American/African (MLA); Other Ethnicity; Residual Categories. For the analyses in this report only the first five are used, as these categories cover the overwhelming number of cases, with only 1.9\% (22/1137) of cases having no Prioritized Ethnicity (see \emph{None}, Table~\autoref{table:none}).

@@

The numbers of cases per category and population sizes for the complete data set from 2007-2014 can be seen in Table~\autoref{table:percap}. Subsequent regression analyses (not shown) also suggested that the MLA category was over-- or underrepresented in per capita rates given the very small sample sizes for this classification (Figure~\autoref{fig:percap}, Table~\autoref{table:percap}), leading to very large standard error in regression analyses. However, there are numerous issues with the data for MLA ethnicity category above and beyond the small population sizes (see Table~\autoref{table:percap}), thus creating issues for estimating the denominator data for this group (University of Otago, personal communication). Thus, we removed this grouping for our subsequent analyses and are left with Asian, European, Maori and Pacific as Prioritized Ethnicities. This left us with 1102/1115 (99\%) of the measles cases with Prioritized Ethnicity recorded from 2007, and 1102/1137 (97\%) of all measles cases recorded since 2007 (Table~\autoref{table:percap}).

\begin{figure}
\begin{center}
\includegraphics{draftfinalreport-010}
\end{center}
\caption{Per capita rates of measles infections for specific age categories (0--2, 3--5, 6--17, 18--24, and 24+ years) and New Zealand deprivation indices (NZDep, 1--5 and 6--10) for 2007--2014, broken down by ethnicity (Table~\autoref{table:percap} for details)}
\label{fig:percap}
\end{figure}

@@

The distribution of the meases cases per category used in the regression analyses are in Figure~\autoref{fig:histcasecat}.
\begin{figure}
\begin{center}
\includegraphics{draftfinalreport-014}
\end{center}
\caption{Distribution of measles cases per category (Table~\autoref{table:percap}) used in the final regression model (Equation~\autoref{eq:reg})}
\label{fig:histcasecat}
\end{figure}

The predicted values from the regression model plotted against the reported cases are shown in Figure~\autoref{fig:predict}, and the residuals are shown in Figure~\autoref{fig:resid}.
\begin{figure}
\begin{center}
\includegraphics{draftfinalreport-015}
\end{center}
\caption{Regression model (Equation~\autoref{eq:reg}) predictions plotted against the case data (Table~\autoref{table:percap})}
\label{fig:predict}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics{draftfinalreport-017}
\end{center}
\caption{Histogram of residuals from the regression model (Equation~\autoref{eq:reg})}
\label{fig:resid}
\end{figure}

The significance of the different predictor variables can be seen in the ANOVA results (Table~\autoref{table:anova}).
\par
\vspace{5mm} %5mm vertical space
\begin{table}
% latex table generated in R 3.0.2 by xtable 1.7-1 package
% Sat Nov 08 03:43:34 2014
\begin{tabular}{lrrrrrr}
  \hline
 & Df & Deviance & Resid. Df & Resid. Dev & F & Pr($>$F) \\ 
  \hline
NULL &  &  & 39 & 1453.62 &  &  \\ 
  Age & 4 & 1304.20 & 35 & 149.43 & 183.58 & 0.0000 \\ 
  Ethnicity & 3 & 20.00 & 32 & 129.43 & 3.75 & 0.0285 \\ 
  NZDep & 1 & 10.70 & 31 & 118.74 & 6.02 & 0.0239 \\ 
  Age:Ethnicity & 12 & 81.91 & 19 & 36.83 & 3.84 & 0.0045 \\ 
   \hline
\end{tabular}\caption{Significance of different predictor variables for measles risk factors from 2007-2014, estimated using Equation~\autoref{eq:reg} and a quasipoisson error structure}
\label{table:anova}
\end{table}
\vspace{5mm} %5mm vertical space

A summary of the regression model (Equation~\autoref{eq:reg}) with the individual effects and the statistical support for the estimated coefficients can be seen in Table~\autoref{table:regression}.

\vspace{5mm} %5mm vertical space
\begin{table}
% latex table generated in R 3.0.2 by xtable 1.7-1 package
% Sat Nov 08 03:43:34 2014
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
  \hline
(Intercept) & -6.4674 & 0.1148 & -56.32 & 0.0000 \\ 
  Age3-5 & -0.9129 & 0.2054 & -4.44 & 0.0003 \\ 
  Age6-17 & -0.7618 & 0.1343 & -5.67 & 0.0000 \\ 
  Age18-24 & -1.5050 & 0.1937 & -7.77 & 0.0000 \\ 
  Age25+ & -2.9507 & 0.1608 & -18.35 & 0.0000 \\ 
  EthnicityAsian & 0.0530 & 0.3250 & 0.16 & 0.8721 \\ 
  EthnicityMaori & 0.2124 & 0.1994 & 1.07 & 0.2999 \\ 
  EthnicityPacific & 1.1603 & 0.2043 & 5.68 & 0.0000 \\ 
  NZDep6-10 & -0.2119 & 0.0855 & -2.48 & 0.0227 \\ 
  Age3-5:EthnicityAsian & -2.0315 & 1.3827 & -1.47 & 0.1581 \\ 
  Age6-17:EthnicityAsian & -1.0074 & 0.4717 & -2.14 & 0.0459 \\ 
  Age18-24:EthnicityAsian & -0.8311 & 0.5941 & -1.40 & 0.1779 \\ 
  Age25+:EthnicityAsian & 0.4211 & 0.4433 & 0.95 & 0.3541 \\ 
  Age3-5:EthnicityMaori & -0.3864 & 0.4096 & -0.94 & 0.3573 \\ 
  Age6-17:EthnicityMaori & -0.0855 & 0.2469 & -0.35 & 0.7328 \\ 
  Age18-24:EthnicityMaori & -0.5828 & 0.4484 & -1.30 & 0.2092 \\ 
  Age25+:EthnicityMaori & -1.0482 & 0.5242 & -2.00 & 0.0600 \\ 
  Age3-5:EthnicityPacific & -2.1316 & 0.8139 & -2.62 & 0.0169 \\ 
  Age6-17:EthnicityPacific & -1.4541 & 0.3347 & -4.35 & 0.0003 \\ 
  Age18-24:EthnicityPacific & -0.9827 & 0.5129 & -1.92 & 0.0705 \\ 
  Age25+:EthnicityPacific & -0.6127 & 0.4367 & -1.40 & 0.1767 \\ 
   \hline
\end{tabular}\caption{Summary of the regression model results for measles risk factors for cases from 2007-2014}
\label{table:regression}
\end{table}
\vspace{5mm} %5mm vertical space

\par Apart from over-representation of some MLA categories discussed above and not included here, the results of the regression model suggest that age is a strong predictor of being a measles case. Indeed, all age categories are significantly less likely to be measles cases compared to 0-2 year olds, and the likelihood generally decreases with age (see also Figure~\autoref{fig:ageinyears}).

The majority of measles cases are among people of European ancestry (Table~\autoref{table:percap} and Figure~\autoref{fig:ageinyears}), however, people of Pacific origin are also over-represented as measles cases on a per capita basis ($\beta$ = 1.16, standard error (SE) = 0.2, p--value < 0.0001). Those people in NZDep levels 6-10 are under-represented ($\beta$ = -0.21, SE = 0.085, p--value = 0.02), and there are some other age:ethnicity classes that are significantly less represented in the data compared to Europeans in those ages classes, particularly in the 6--17 age classes. In later measles outbreaks (since 2007) there has been a shift in the distribution of ages infected. The very young (<2 years old) are still most likely to be infected, but school aged children and older teenagers are more likely to be represented then the under tens (Figures~\autoref{fig:ageinyears} and ~\autoref{fig:ageandvac}, Table~\autoref{table:sero}, and section~\autoref{sub:popim}). This pattern suggests that improving vaccination coverage in the young is reducing the burden of measles in those age categories (Table~\autoref{table:percap}). Interestingly the regression results suggest risk of measles in the 6--17 year age category was greater for Europeans and Maori.

@@

The majority of the na\"{\i}ve among the general New Zealand population are in their first years of life (Figure~\autoref{fig:naive}). However, the distribution of na\"{\i}ve at a national level shows that the recent MMR vaccination schemes are reducing the proportions of na\"{\i}ve population in the 3--5 year old age group to greater levels than in older young people  (Figure~\autoref{fig:naive}). Plotting the breakdown of these figures by DHB clearly shows that the greatest numbers of na\"{\i}ve people are in DHBs with larger urban areas (Figure~\autoref{fig:naive_map}). The distribution of the numbers of na\"{\i}ve and the total na\"{\i}ve populations per DHB, assuming national immunisation and immunity rates are representative, are given in the following figures:
\begin{itemize}[noitemsep,nolistsep]
\item Northland: Figure~\autoref{fig:Northland}
\item Waitemata: Figure~\autoref{fig:Waitemata}
\item Auckland: Figure~\autoref{fig:Auckland}
\item Counties Manukau: Figure~\autoref{fig:Counties_Manukau}
\item Waikato: Figure~\autoref{fig:Waikato}
\item Lakes: Figure~\autoref{fig:Lakes}
\item Bay of Plenty: Figure~\autoref{fig:BayofPlenty}
\item Tairawhiti: Figure~\autoref{fig:Tairawhiti}
\item Taranaki: Figure~\autoref{fig:Taranaki}
\item Hawke's Bay: Figure~\autoref{fig:HawkesBay}
\item Whanganui: Figure~\autoref{fig:Whanganui}
\item Mid-Central: Figure~\autoref{fig:Midcentral}
\item Hutt: Figure~\autoref{fig:Hutt}
\item Capital and Coast: Figure~\autoref{fig:CapitalandCoast}
\item Wairarapa: Figure~\autoref{fig:Wairarapa}
\item Nelson Marlborough: Figure~\autoref{fig:NelsonMarlborough}
\item West Coast: Figure~\autoref{fig:WestCoast}
\item Canterbury: Figure~\autoref{fig:Canterbury}
\item South Canterbury: Figure~\autoref{fig:SouthCanterbury}
\item Southern: Figure~\autoref{fig:Southern}
\end{itemize}

\begin{figure}
\begin{center}
\includegraphics{draftfinalreport-023}
\end{center}
\caption{New Zealand population by age and estimated numbers of na\"{\i}ve people in each age class using national immunity data (Table~\autoref{table:sero})}
\label{fig:naive}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=1.2\textwidth]{naive_map.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:naive_map}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb1.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Northland District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:Northland}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb2.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Waitemata District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:Waitemata}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb3.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Auckland District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:Auckland}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb4.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Counties Manukau District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:Counties_Manukau}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb5.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Waikato District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:Waikato}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb6.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Lakes District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:Lakes}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb7.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Bay of Plenty District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:BayofPlenty}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb8.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Tairawhiti District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:Tairawhiti}
\end{figure}

\begin{figure}[H]
  \begin{center}
    \includegraphics[width=0.6\textwidth]{dhb9.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Taranaki District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:Taranaki}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb10.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Hawke's Bay District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:HawkesBay}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb11.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Whanganui District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:Whanganui}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb12.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Midcentral District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:Midcentral}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb13.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Hutt District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:Hutt}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb14.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Capital and Coast District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:CapitalandCoast}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb15.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Wairarapa District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:Wairarapa}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb16.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Nelson Marlborough District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:NelsonMarlborough}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb17.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, West Coast District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:WestCoast}
\end{figure}

\begin{figure}[H]
   \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb18.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Canterbury District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:Canterbury}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb19.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, South Canterbury District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:SouthCanterbury}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb20.pdf}
     \end{center}
     \caption{Numbers of na\"{\i}ve individuals per age class, Southern District Health Board, using national immunity data (Table~\autoref{table:sero})}
     \label{fig:Southern}
\end{figure}

@@

Though the global incidence of measles is declining, in recent years that decline has slowed (Figure~\autoref{fig:trendincidence}) and immigration rates to New Zealand have risen (Figure~\autoref{fig:travel}). This suggests that the risk of measles importation could increase, though further analyses are require to understand the interaction between these variables.  Of note, however, is the clear seasonality in immigration and travel (Figure~\autoref{fig:travel}).  This seasonality suggests that there may be period of increased risk of measles importation, though again the interactions with seasonal measles transmission from the nations of origin will be an important factor in determining the risk of measles importation. Of interest is the asynchrony, or lack of phase, between peak non-New Zealander immigration and New Zealanders travelling. The Christmas summer period being an exception to this, when there is substantial New Zealander travel and combined these lead to a very sharp peak in travel rates in December (Figure~\autoref{fig:travel}).
